Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-07-06 WholePlotTrajectories.tex
graphs, stats,etc

Six functional traits representing leaf economics (leaves thickness,
toughness, total chlorophyll content and specific leaf area) and wood
economics (wood specific gravity and bark thickness), and life-history
traits (maximum specific height and seed mass) came from the BRIDGE
project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
assessed from a selection of individuals located in nine permanent plots
in French Guiana, including two in Paracou, and comprised 294 species
pertaining to 157 genera. Missing trait values (10\%) were filled using
multivariate imputation by chained equation \citep{Mice2011}.
Imputations were restricted within genus or family when samples were too
scarce, in order to account for the phylogenetic signal. Whenever a
species was not in the dataset, it was attributed a set of trait values
randomly sampled among species of the next higher taxonomic level (same
genus or family). As seed mass information was classified into classes,
no data filling process was applied and analyses were restricted to the
414 botanical species recorded.

@@

The impacts of initial disturbance were tested with the spearman rank
correlation between the extrema of taxonomic and functional metrics
reached over the 30 years and the initial \%AGB removed. The were
besides analysed through the linear correlations between (i) taxonomic
and functional richness, evenness and diversity and (ii) the initial
\%AGB removed at 10, 20 and 30 years after disturbance.

@@

From 1989 (2 years after disturbance) to 2015 (28 years after
disturbance), 828388 individual trees and 591 botanical species
pertaining to 223 genus and 64 families were recorded. For undisturbed
plots, taxonomic Richness and Evenness remained stable over the 30 years
monitored. In disturbed communities, after low disturbance intensity the
taxonomic richness increased, reaching a maximum gain of 14 botanical
genera (plot 3 from treatment 2). After intense disturbance the
taxonomic richness followed a more complex trajectory, decreasing for
ten years after disturbance before recovering to pre-disturbance values.
In all disturbed plots the taxonomic evenness first increased until a
maximum reached after around 20 years. This maximum was positively
correlated to the disturbance intensity
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). The evenness then stabilized except
for two T3 plots (plots 8 and 12) for which evenness still increased.

@@

\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
\end{figure*}

@@

The impact of disturbance intensity was examined through the linear
correlation between (i) the \%AGB removed and (ii) taxonomic richness,
evenness and functional diversity after 10, 20 and 30 years
\ref{fig:IDHplot}. The correlation with disturbance intensity was weak
for te taxonomic richness (\(-0.09<R^2<-0.06\)), for the taxonomic
evenness (\(R^2<0.25\)) while highly significant for functional
diversity (\(0.60<R^2<0.75\)) whatever the time since disturbance.

@@

\caption{Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals. Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
\end{figure*}

@@

While both taxonomic and functional composition remained stable in
undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
marked and consistent trajectories over post-disturbance time. in
disturbed communities, these compositional changes corresponded to
shifts towards species with more acquisitive functional strategies, from
communities with high average WD to high average SLA and chlorophyll
content (see appendix I). For functional composition, this translated
into cyclic compositional changes with an unachieved recovery of the
initial composition (Figure \ref{fig:NMDSplans}). The maximum
dissimilarity with the initial state was positively correlated to the
disturbance intensity for both taxonomic and functional composition
(\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively). The maximum value
was reached around 26 years after disturbance for taxonomic composition
and 22 years for functional composition.

@@

Maximum height at adult stage (\emph{Hmax}), leaf toughness
(\emph{L\_toughness}) and wood density (\emph{WD}) first decreased and
then slightly increased but remained significantly lower than their
initial value (Figure \ref{fig:CWM}). On the other side, Bark thickness
(\emph{Bark\_thick}) and specific leaf area (\emph{SLA}) increased and
while \emph{Bark\_thick} remained substantially high after 30 years,
\emph{SLA} had almost recovered its initial value. For all traits, the
maximum difference to initial value was correlated to the disturbance
intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
\(\rho_{spearman}^{L_{chloro}}=0.45\),
\(\rho_{spearman}^{L_{toughness}}=-0.43\),
\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WD}=-0.78\),
\(\rho_{spearman}^{Bark-thickness}=0.88\),
\(\rho_{spearman}^{Hmax}=-0.48\)).

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
\end{figure*}

@@

All disturbed plots had lower functional redundancy than control plots
and followed similar hump-shaped trajectories (@ref(fig:RedFun\_rest)).
The maximum redundancy loss was positively correlated with the
disturbance intensity (\(\rho_{spearman}=0.50\)) and the initial value
had not recovered for any disturbed communities after 30 years.

@@

\caption{TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
\end{figure}

@@

Community functional composition followed similar trajectories that were
not so distinct in the functional space\ldots{}\ldots{}. Because the
pre-disturbance survivor trees mirror the initial communities
\citep{Herault2018}, the functional changes relied upon recruited trees
and the enhanced growth and survival of previously infrequent species
and functional types. Disturbance vacated environmental niches of high
light, space and nutrient availability, filled by competitive pioneers
that became dominant and determined the functional characteristics of
the community \citep{Grime1998}. Before disturbance and over communities
recovery time though, these species would be excluded by long-lived,
more resistant and shade tolerant species. The recovery of communities
functional characteristics, that mostly rely on dominant species
according to the ``vegetation quantity effect'' \citep{Grime1998},
suggested a similar recovery of the dominant species \citep{Molino2001}.
significant functional shifts towards resource-acquisitive strategies
(sharp increase in the SLA, leaf thickness and bark thickness and
decrease in wood density, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}.

>>>>

Six functional traits representing leaf economics (leaves thickness,
toughness, total chlorophyll content and specific leaf area) and stem
economics (wood specific gravity and bark thickness), and life-history
traits (maximum specific height and seed mass) came from the BRIDGE
project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
assessed from a selection of individuals located in nine permanent plots
in French Guiana, including two in Paracou, and comprised 294 species
pertaining to 157 genera. Missing trait values (10\%) were filled using
multivariate imputation by chained equation \citep{Mice2011}.
Imputations were restricted within genus or family when samples were too
scarce, in order to account for the phylogenetic signal. Whenever a
species was not in the dataset, it was attributed a set of trait values
randomly sampled among species of the next higher taxonomic level (same
genus or family). As seed mass information was classified into classes,
no data filling process was applied and analyses were restricted to the
414 botanical species recorded.

@@

The impacts of initial disturbance were tested with the spearman rank
correlation between the extrema of taxonomic and functional metrics
reached over the 30 years and the initial \%AGB removed. They were
besides analysed through polynomial regression between (i) taxonomic and
functional richness, evenness and diversity and (ii) the initial \%AGB
removed at 10, 20 and 30 years after disturbance.

@@

From 1989 (2 years after disturbance) to 2015 (28 years after
disturbance), 828 388 individual trees and 591 botanical species
pertaining to 223 genus and 64 families were recorded. For undisturbed
plots, taxonomic Richness and Evenness remained stable over the 30 years
monitored. In disturbed communities, after low disturbance intensity the
taxonomic richness increased, reaching a maximum gain of 14 botanical
genera (plot 3 from treatment 2). After intense disturbance the
taxonomic richness followed a more complex trajectory, decreasing for
ten years after disturbance before recovering to pre-disturbance values.
The maximum richness loss or gain after disturbance was positively
correlated to the disturbance intensity
(\(\rho_{spearman}^{Richness}=0.50\)). In all disturbed plots the
taxonomic evenness first increased until a maximum reached after around
20 years. This maximum was positively correlated to the disturbance
intensity (\(\rho_{spearman}^{Evenness}=0.77\)). The evenness then
stabilized except for two T3 plots (plots 8 and 12) for which evenness
kept increasing.

@@

\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness and \textbf{(c)} functional diversity. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
\end{figure*}

@@

Among polynomial regressions between (i) the \%AGB removed and (ii)
taxonomic richness, evenness and functional diversity after 10, 20 and
30 years the second degree better predicted the hump-shaped curve of the
disturbance impact in function of the intensity \ref{fig:IDHplot}. The
correlation with disturbance intensity was weak for the taxonomic
richness (\(R^2<0.37\)) and taxonomic evenness (\(R^2<0.43\)) but highly
significant for functional diversity (\(0.60<R^2<0.81\)).

@@

\caption{Relationship between the initial \%AGB removed and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)} and functional diversity \textbf{(c)} at 10, 20 and 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
\end{figure*}

@@

While both taxonomic and functional composition remained stable in
undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
marked and consistent trajectories over post-disturbance time. in
disturbed communities, these compositional changes corresponded to
shifts towards species with more acquisitive functional strategies, from
communities with high average WSG to high average SLA and chlorophyll
content (see appendix I). For functional composition, this translated
into cyclic compositional changes with an unachieved recovery of the
initial composition (Figure \ref{fig:NMDSplans}). The maximum
dissimilarity with the initial state was positively correlated to the
disturbance intensity for both taxonomic and functional composition
(\(\rho_{spearman}^{taxonomic}=0.87\) and
\(\rho_{spearman}^{functional}=0.90\) respectively). The maximum value
was reached around 26 years after disturbance for taxonomic composition
and 22 years for functional composition.

@@

Maximum height at adult stage (\emph{Hmax}), leaf toughness
(\emph{L\_toughness}) and wood specific gravity (\emph{WSG}) first
decreased and then slightly increased but remained significantly lower
than their initial value (Figure \ref{fig:CWM}). On the other side, Bark
thickness (\emph{Bark\_thick}) and specific leaf area (\emph{SLA})
increased and while \emph{Bark\_thick} remained substantially high after
30 years, \emph{SLA} had almost recovered its initial value. For all
traits, the maximum difference to initial value was correlated to the
disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.76\),
\(\rho_{spearman}^{L_{chloro}}=0.60\),
\(\rho_{spearman}^{L_{toughness}}=-0.53\),
\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WSG}=-0.75\),
\(\rho_{spearman}^{Bark-thickness}=0.71\),
\(\rho_{spearman}^{Hmax}=-0.40\)).

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WSG}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
\end{figure*}

@@

All disturbed plots had lower functional redundancy than control plots
and followed similar hump-shaped trajectories (@ref(fig:RedFun\_rest)).
The maximum redundancy loss was positively correlated with the
disturbance intensity (\(\rho_{spearman}=0.47\)) and the initial value
had not recovered for any disturbed communities after 30 years.

@@

\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
\end{figure}

@@

Community functional composition followed similar trajectories that were
not so distinct in the functional space\ldots{}\ldots{}. Because the
pre-disturbance survivor trees mirror the initial communities
\citep{Herault2018}, the functional changes relied upon recruited trees
and the enhanced growth and survival of previously infrequent species
and functional types. Disturbance vacated environmental niches of high
light, space and nutrient availability, filled by competitive pioneers
that became dominant and determined the functional characteristics of
the community \citep{Grime1998}. Before disturbance and over communities
recovery time though, these species would be excluded by long-lived,
more resistant and shade tolerant species. The recovery of communities
functional characteristics, that mostly rely on dominant species
according to the ``vegetation quantity effect'' \citep{Grime1998},
suggested a similar recovery of the dominant species \citep{Molino2001}.
significant functional shifts towards resource-acquisitive strategies
(sharp increase in the SLA, leaf thickness and bark thickness and
decrease in wood specific gravity, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}.
